`r params$doc_title`

### 2022-11-15 scripts/MSc.Rmd
update be4 nightshift

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | lorenzo@imbi.uni-heidelberg.de
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    code_folding: hide
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
#bibliography: ../references/project01.bib
---

@@

```{r simulatesimulatesimulate}
## try starting with n=50.000 -> selected 9 IVs
set.seed(1234)

## initialize
nsim <- 100
se_update_BMI <- c()
se_update_GBC <- c()

est <- tibble(no_sim = 1:nsim)
#est <- matrix(NA, 
#              nrow = nsim, 
#              ncol = 2*length(n)  # theta_sim and theta_se_sim
#              ) 
#colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))

## START SIM
T0 = proc.time()[3]
for (i in 1:length(colnames_pval)) {
  T0 = proc.time()[3]
  threshold <- 5*10^-6
  numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
  if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
    next
  }

  # FILTER DATA
  sim_dat <- my_data_harm %>% 
    filter(!!sym(colnames_pval[i]) < threshold) %>% 
    select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)

  # get simulated se_{sample size}
  for (j in 1:nrow(sim_dat)) {
    # as done before in section "simulating on empirical data"
    se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
    se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
  }

  sim_dat <- sim_dat %>%
      bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))

  # simulate beta values from normal_distribution
  for (k in 1:nsim) {
    sim_dat <- sim_dat %>% 
      mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
      mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))

    # PERFORM MR_analysis
    mr.obj = MendelianRandomization::mr_input(
          bx = sim_dat$beta.bmi_sim_norm, 
          bxse = sim_dat$se_update_BMI, 
          # outcome
          by = sim_dat$beta.gbc_sim_norm, 
          byse = sim_dat$se_update_GBC, 
          snps = sim_dat$SNP,
          exposure = "Body mass index",
          outcome = "Gallbladder Cancer"
        ) 
    # CONMIX (as reference)
    res = MendelianRandomization::mr_conmix(mr.obj)
    colname_theta <- paste0("theta_sim_", n[i])
    colname_theta_se <- paste0("theta_se_sim_", n[i])
    ## store results
    est[k, colname_theta] <- res$Estimate
    CIlength = res$CIUpper-res$CILower
    if (length(CIlength)>1) print("conmix multimodal")
    est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
    rm(res)
  }
  print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
}
T1 = proc.time()[3]
timediff = T1-T0
```

>>>>

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    code_folding: hide
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
bibliography: ../references/references.bib
---

@@

```{r simulatesimulatesimulate, warning=FALSE}
## try starting with n=50.000 -> selected 9 IVs
# set.seed(1234)
# 
# ## initialize
# nsim <- 25
# est <- tibble(no_sim = 1:nsim)
# #est <- matrix(NA, 
# #              nrow = nsim, 
# #              ncol = 2*length(n)  # theta_sim and theta_se_sim
# #              ) 
# #colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
# 
# ## START SIM
# T0 = proc.time()[3]
# for (i in 1:length(colnames_pval)) {
#   threshold <- 5*10^-6
#   se_update_BMI <- c()
#   se_update_GBC <- c()
#   
#   numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
#   
#   if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
#     next
#   }
#   
#   # FILTER DATA
#   sim_dat <- my_data_harm %>% 
#     filter(!!sym(colnames_pval[i]) < threshold) %>% 
#     select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
#   
#   # get simulated se_{sample size}
#   for (j in 1:nrow(sim_dat)) {
#     # as done before in section "simulating on empirical data"
#     se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
#     se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
#   }
#   
#   sim_dat <- sim_dat %>%
#       bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
#   
#   # simulate beta values from normal_distribution
#   for (k in 1:nsim) {
#     sim_dat <- sim_dat %>% 
#       mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
#       mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))
#     
#     # PERFORM MR_analysis
#     mr.obj = MendelianRandomization::mr_input(
#           bx = sim_dat$beta.bmi_sim_norm, 
#           bxse = sim_dat$se_update_BMI, 
#           # outcome
#           by = sim_dat$beta.gbc_sim_norm, 
#           byse = sim_dat$se_update_GBC, 
#           snps = sim_dat$SNP,
#           exposure = "Body mass index",
#           outcome = "Gallbladder Cancer"
#         ) 
#     # CONMIX (as reference)
#     res = MendelianRandomization::mr_conmix(mr.obj)
#     colname_theta <- paste0("theta_sim_", n[i])
#     colname_theta_se <- paste0("theta_se_sim_", n[i])
#     ## store results
#     est[k, colname_theta] <- res$Estimate
#     CIlength = res$CIUpper-res$CILower
#     if (length(CIlength)>1) print("conmix multimodal")
#     est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
#     rm(res)
#   }
#   rm(se_update_BMI, se_update_GBC)
#   print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
# }
# 
# 
# T1 = proc.time()[3]
# timediff = T1-T0
# 
# 
# saveRDS(est, paste0(Sys.Date(), "_est.rds"))
```
